Quantitative Benchmarking for Automatically Generated Conjectures

### 2018-01-15 TEBenchmarkPaper/article.tex
Fleshing out table more
More table justification

% For interestingness table
\newcommand{\iE}{\textbf{E}}
\newcommand{\iN}{\textbf{N}}
\newcommand{\iS}{\textbf{S}}
\newcommand{\iA}{\textbf{A}}
\newcommand{\iC}{\textbf{C}}
\newcommand{\iU}{\textbf{U}}
\newcommand{\tIFF}{if-and-only-if}
\newcommand{\tNE}{non-exists}
\newcommand{\tIMP}{implies}
\newcommand{\tRow}[1]{#1 \\ \hline}

@@

\begin{table}
  \centering
  \begin{tabular}{ |l|l|c|c|c|c|c|c| }
    \hline
    \multirow{2}{*}{\textbf{Program}}                     &
    \multirow{2}{*}{\textbf{Conjecture Types}}            &
    \multicolumn{6}{c}{\textbf{Interestingness Measures}} \\ \hhline{~~------}
    \tRow{            &                    & \iE & \iN & \iS & \iA & \iC & \iU}
    \tRow{AM          & \tIFF, \tIMP, \tNE &   X &   X &   X &   X &   X &   X}
    \tRow{GT          & \tIFF, \tIMP, \tNE &   X &   X &   X &   X &   X &   X}
    \tRow{Graffiti    & inequalities       &   X &   X &   X &     &   X &   X}
    \tRow{Bagai et al & \tNE               &     &   X &     &   X &   X &    }
    \tRow{HR          & \tIFF, \tIMP, \tNE &   X &   X &   X &   X &   X &   X}
    \tRow{QuickSpec   & equalities         &   X &   X &     &     &   X &    }
    \tRow{IsaCoSy     & equalities         &     &   X &     &     &   X &    }
    \tRow{IsaScheme   & equalities         &     &     &     &     &   X &    }
  \end{tabular}
  \caption{Classification of ATE systems from \cite{colton2000notion}, extended
    to those compared in \cite{claessen2013automating} (QuickSpec is the
    conjecture generation component of HipSpec). The interestingness measures
    are \iE{}mpirical plausibility, \iN{}ovelty, \iS{}urprisingness,
    \iA{}pplicability, \iC{}omprehensibility (low complexity) and \iU{}tility.}
  \label{table:colton}
\end{table}

We summarise the criteria in Table 1 and state which were used in key
ATF/ATE systems, where Colton identified usage in AM, GT, Graffiti and
Bagai et al, and his own ATF system HR, and we identify usage in three
more recent ATE systems QuickSpec, IsaCoSy and IsaScheme. These
criteria, especially empirical plausibility and comprehensibility, are
not only used for analysing results after the fact, but instead may
form a core part of an algorithm's design decisions. For example, the
designers of IsaCoSy consider simple substitutions of existing
conjectures (i.e. those with low novelty) to be uninteresting, and
hence their system includes a constraint solving component which
avoids generating such statements entirely.

With such ambiguous and varied goals, approaches and assessment
criteria it is difficult to compare ATE systems in a quantitative way,
and hence to form some measure of ``progress'' for the field.

>>>>

% For interestingness table
\newcommand{\iE}{\textbf{E}}
\newcommand{\iN}{\textbf{N}}
\newcommand{\iS}{\textbf{S}}
\newcommand{\iA}{\textbf{A}}
\newcommand{\iC}{\textbf{C}}
\newcommand{\iU}{\textbf{U}}
\newcommand{\tIFF}{if-and-only-if}
\newcommand{\tNE}{non-exists}
\newcommand{\tIMP}{implies}
\newcommand{\tEQ}{equations}
\newcommand{\tRow}[1]{#1 \\ \hline}

@@

\begin{table}
  \centering
  \begin{tabular}{ |l|l|c|c|c|c|c|c| }
    \hline
    \multirow{2}{*}{\textbf{Program}}                     &
    \multirow{2}{*}{\textbf{Conjecture Types}}            &
    \multicolumn{6}{c}{\textbf{Interestingness Measures}} \\ \hhline{~~------}
    \tRow{            &                    & \iE & \iN & \iS & \iA & \iC & \iU}
    \tRow{AM          & \tIFF, \tIMP, \tNE &   X &   X &   X &   X &   X &   X}
    \tRow{GT          & \tIFF, \tIMP, \tNE &   X &   X &   X &   X &   X &   X}
    \tRow{Graffiti    & inequalities       &   X &   X &   X &     &   X &   X}
    \tRow{Bagai et al & \tNE               &     &   X &     &   X &   X &    }
    \tRow{HR          & \tIFF, \tIMP, \tNE &   X &   X &   X &   X &   X &   X}
    \tRow{QuickSpec   & \tEQ               &   X &   X &     &     &   X &    }
    \tRow{Speculate   & conditional \tEQ   &   X &   X &     &   A &   X &    }
    \tRow{IsaCoSy     & \tEQ               &   X &   X &     &     &   X &   X}
    \tRow{IsaScheme   & \tEQ               &   X &   X &     &     &   X &   X}
  \end{tabular}
  \caption{Classification of ATE systems from \cite{colton2000notion}, extended
    to those compared in \cite{claessen2013automating} (QuickSpec is the
    conjecture generation component of HipSpec). The interestingness measures
    are \iE{}mpirical plausibility, \iN{}ovelty, \iS{}urprisingness,
    \iA{}pplicability, \iC{}omprehensibility (low complexity) and \iU{}tility.}
  \label{table:colton}
\end{table}

% JUSTIFICATIONS
%
% IsaCoSy uses counter-example checking; this ensures empirical plausibility
% IsaCoSy uses constraints to avoid special-cases; this ensures novelty
% IsaCoSy uses utility, since there are some hard-coded patterns which are
% looked for
%
% Speculate uses a form of unification to ensure novelty, making sure one
% equation is not a special-case of another
% Speculate uses LeanCheck to enumerate values, looking for counterexamples
% Speculate uses testing, to ensure Empirical Plausibility
%
% QuickSpec uses QuickCheck to ensure empirical plausibility
% QuickSpec uses a congruence closure algorithm to ensure novelty
%
% IsaScheme uses utility, since it determines the "quality" of a definition
% based on how many theorems it appears in?
% IsaScheme uses novelty, using an equational rewrite system (Knuth-Bendix
% completion) to remove redundancies
% IsaScheme uses utility, since it focuses on the terms and definitions of a
% user's theory. However, don't all of them?
% IsaScheme uses counterexample checking for empirical plausibility.

% ONLY APPLIES TO PRECONDITIONS
% APPLICABILITY: ensure a conjecture holds in many models (or exactly zero, in
% the case of Bagai et al). Does the testing-based approach of QuickSpec and
% Speculate ensure that there's a model? After all, these are concrete values
% rather than e.g. implications of some abstract specification.

We summarise the criteria in Table 1 and state which were used in key
ATF/ATE systems. The entries for AM, GT, Graffiti, Bagai et al. and HR are based
on this survey (the latter is Colton's own system). We extend this analysis to
some more recent systems in the remaining rows (based on our understanding of
their function): QuickSpec, Speculate, IsaCoSy and IsaScheme.

All four of these systems use counterexample checking and a small-to-large
search order, favouring empirical plausibile and comprehensibile output. IsaCoSy
ensures novelty using a constraint system, such that no generated expression can
be simplified using a previously-found equation. The other systems use post-hoc
filters based on unification and rewriting (based on the Knuth-Bendix completion
algorithm) to discard expressions or equations which are special-cases of some
other, more general result.

IsaCoSy is able to look for commonly desirable patterns, such as commutativity
and associativity, which increases the utility of its output, whilst IsaScheme
is completely based around such pattern-instantiation.

This diversity of approaches makes it difficult to compare the existing
evaluations of these systems directly. In particular, evaluation methods created
for one system might not make sense for another, due to assumptions made about
the algorithms' operation. For example, the novelty filters used by QuickSpec
and Speculate are applied to the whole output set, guaranteeing that no
special-cases will appear; yet we cannot assume this in the case of IsaCoSy,
since its constraint solver allows special cases if they're found \emph{before}
the general case (it does not go back and discard previous results).

With such ambiguous and varied goals, approaches and assessment criteria it is
difficult to compare ATE systems in a quantitative way, and hence to form some
measure of ``progress'' for the field. Our proposed benchmarking methodology
attempts to remedy this by setting a clear target, hopefully in line with the
myriad goals of existing researchers, whilst being generally applicable to
systems of this type.
